---
title: "Time Series Analysis Tutorial"
author: "Henry Bernreuter"
date: "2024-10-03"
output: html_document
---
# Introduction:
## In this tutorial, we will perform time series analysis on a dataset containing search term analytics. We will load the data, create a time series object, partition the data, generate forecasts using different models, and evaluate their accuracy.

## Load Data and Create Time Series ObjectFirst, we will load the data and create a time series object.

```{r}
knitr::opts_chunk$set(echo = TRUE)

# Clear Environment
rm(list = ls())


# Load data
Search.data <- read.csv("Kaggle_SearchTerm.csv")

# Create time series object
Search.ts <- ts(Search.data$analytics, start = c(2014, 51), end = c(2019, 50), freq = 52)

# Plot the time series
plot(Search.ts, xlab = "Time", ylab = "Analytics")
```

# Partition the Data next, we will partition the data into training and validation sets.
```{r setup, include=FALSE}


# Partition the data
nValid <- 100
nTrain <- length(Search.ts) - nValid

train.ts <- window(Search.ts, start = c(2014, 51), end = c(2018, 2))
valid.ts <- window(Search.ts, start = c(2018, 3), end = c(2019, 50))

```

# Generate Naive and Seasonal Naive ForecastsWe will generate naive and seasonal naive forecasts and plot them.
```{r setup, include=FALSE}
library(forecast)

# Generate naive and seasonal naive forecasts
naive.pred <- naive(train.ts, h = nValid)
snaive.pred <- snaive(train.ts, h = nValid)

# Plot the forecasts
plot(train.ts, ylab = "Analytics", xlab = "Time", bty = "l", xlim = c(2014.75, 2019), main = "")
lines(naive.pred$mean, lwd = 2, col = "blue", lty = 1)
lines(snaive.pred$mean, lwd = 2, col = "blue", lty = 1)
lines(valid.ts, col = "grey20", lty = 3)

# Evaluate the accuracy
accuracy(naive.pred, valid.ts)
accuracy(snaive.pred, valid.ts)
```
# Linear Trend ModelWe will fit a linear trend model and evaluate its accuracy.
```{r setup, include=FALSE}
# Fit a linear trend model
train.lm <- tslm(train.ts ~ trend)
summary(train.lm)

# Generate forecasts
train.lm.pred <- forecast(train.lm, h = nValid, level = 0)

# Plot the forecasts
plot(train.lm.pred, ylab = "Analytics", xlab = "Time", bty = "l", xaxt = "n", xlim = c(2014.75, 2019), main = "", flty = 2)
lines(train.lm.pred$fitted, lwd = 2, col = "blue")
lines(valid.ts)

# Evaluate the accuracy
accuracy(train.lm.pred, valid.ts)
```

# Quadratic Trend ModelWe will fit a quadratic trend model and evaluate its accuracy.

```{r setup, include=FALSE}

# Fit a quadratic trend model
train.lm.poly.trend <- tslm(train.ts ~ trend + I(trend^2))
summary(train.lm.poly.trend)

# Generate forecasts
train.lm.poly.trend.pred <- forecast(train.lm.poly.trend, h = nValid, level = 0)

# Plot the forecasts
plot(train.lm.poly.trend.pred, ylab = "Analytics", xlab = "Time", bty = "l", xlim = c(2014.75, 2019), main = "")
lines(train.lm.poly.trend$fitted, lwd = 2)
lines(valid.ts)

# Evaluate the accuracy
accuracy(train.lm.poly.trend.pred, valid.ts)
```
# Seasonality ModelWe will fit a seasonality model and evaluate its accuracy.
```{r setup, include=FALSE}
# Fit a seasonality model
train.lm.season <- tslm(train.ts ~ season)
summary(train.lm.season)

# Generate forecasts
train.lm.season.pred <- forecast(train.lm.season, h = nValid, level = 0)

# Evaluate the accuracy
accuracy(train.lm.season.pred, valid.ts)
```

```{r setup, include=FALSE}
# Plot the forecasts
plot(train.lm.season.pred, ylab = "Analytics", xlab = "Time", bty = "l", xlim = c(2014.75, 2019), main = "", flty = 2)
lines(train.lm.poly.trend$fitted, lwd = 2)
lines(valid.ts)
```

# ConclusionIn this tutorial, we performed time series analysis on a dataset containing search term analytics. We generated forecasts using naive, seasonal naive, linear trend, quadratic trend, and seasonality models, and evaluated their accuracy.

